---
title: "Week 4 Data Analysis Script"
author: "Ken"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(reshape2)
```


## overview 

this report represent preliminary data analyis for thr week 3 example data. Specifically, this will report basic descriptive statistics as well as report tests of the following predictions:

1. people will feel more guilt when:
  + 1.1 they caused the shock (vs. their partner)
  + 1.2 their partner was morally good (vs. bad)
  + 1.3 the shock was stromger (high > medium > low)
2. women will report more guilt than men
3. agreeable people will report more guilt
4. neurotic people will report more guilt 

and

5. People will be more generous when:
  + 5.1 they caused the shock (vs. their partner)
  + 5.2 the partner was morally good (vs. bad)
  + 5.3 the shock was stronger (high > medium > low)
6. women will be more generous than men 
7. agreeable people will be more generous 
8. neurotic people will be more generous 



``` {r load}

#Load the data
data<-read.csv("Week 3 /Processed Data/Week 3 Data PROCESSED.csv")

#recode variables
data$PIN<-as.factor(data$PIN)
data$sex<-as.factor(data$sex)
data$shock<-as.factor(data$shock)
data$shockCause<-as.factor(data$shockCause)
data$pMoral<-as.factor(data$pMoral)
```




## Descriptives 

Descriptive statistics for guilt 
```{r desGuilt}

summary(data$guilt)
c(summary(data$guilt),"SD"=sd(data$guilt,na.rm=T))



```


Descriptive statistics for generosity 
```{r desGenerosity}

#Generosity before the shock
c(summary(data$preShare),"SD"=sd(data$preShare,na.rm=T))

#Generosity after the shock
c(summary(data$postShare),"SD"=sd(data$postShare,na.rm=T))
```

Descriptive statistics for personality 
```{r desPers}

#Descriptive for agreeableness
c(summary(data$agre),"SD"=sd(data$agre,na.rm=T))

#Descriptive for neuroticism
c(summary(data$neuro),"SD"=sd(data$neuro,na.rm=T))
``` 



## Guilt Analyses

### Prediction 1.1 
Prediction 1.1 was supported. People felt more guilty when they caused the shock than when their partner did: 
```{r p1.1}

#Run the t-test for prediction 1.1 
p1.1<-t.test(data$guilt~data$shockCause)

#output the result
p1.1
```




### prediction 1.2 

Prediction 1.2 was supported. People felt more guilty when their partner was morally good:
```{r p1.2}

p1.2<-t.test(data$guilt~data$pMoral)


#Output the result 
p1.2 

```



### prediction 1.3 

Prediction 1.3 was supported. People felt more guilty when their partner recieved a stronger shock: 
```{r p1.3}

p1.3<-lm(guilt~shock,data=data)


#Output the result 
anova(p1.3)

#Output the group means
tapply(data$guilt,data$shock,function(x) mean(x,na.rm=T))

```


### Prediction 2 

Prediction 2 was not supported. women did not report more guilt than men:
```{r p2}

p2<-t.test(data$guilt~data$sex)


#output the result 
p2 
```



### Prediction 3 
Prediction 3 was supported. agreeable people were more guilty. 
```{r p3}

#runs the correlation test for prediction 3 
p3<-cor.test(data$guilt,data$agre)

#output the results 
p3
```




### Prediction 4 
Prediction 4 was not supported. Neurotic people were not more guilty 
```{r p4}


#Runs the correlation test 
p4<-cor.test(data$guilt,data$neuro)


#Output the results 
p4
```



## Generosity Analysis 

``` {r melt}
#Melt our data from wide format to long format
dataMelt<-melt(data,
               measure.vars=c("preShare","postShare"),
               variable.name=c("prePost"),
               value.name=c("share")
               )

```




### Prediction 5.1 
Prediction 5.1 was supported. People were more generous when they caused the shock and were actually less generous when the partner caused the shock: 
```{r p5.1}

p5.1<-aov(share~prePost*shockCause+Error(PIN),data=dataMelt)


#Output the results:
summary(p5.1)

#Output the means:
tapply(dataMelt$share,dataMelt[,c(24,21)],function(x) mean(x,na.rm=T))
```


